\documentclass[a4paper,12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[left=0.5in,right=0.5in,top=1in,bottom=1in]{geometry}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{pgfplots,graphicx,calc,changepage}
\pgfplotsset{compat=newest}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage[colorlinks = true, linkcolor = blue]{hyperref}

% Syntax highlighting
\usepackage{listings}
\usepackage{xcolor}

\definecolor{codegreen}{rgb}{0.40,0.62,0.07}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codeblue}{rgb}{0.09,0.57,0.73}
\definecolor{backcolour}{rgb}{1,1,1}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codeblue},
    basicstyle=\ttfamily\small,
    breaklines=true,                     
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,
    showstringspaces=false,
    showtabs=false,                  
    tabsize=4
}

\lstset{style=mystyle}

\newcommand{\nats}{\mathbb{N}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\rats}{\mathbb{Q}}
\newcommand{\ints}{\mathbb{Z}}
\newcommand{\comps}{\mathbb{C}}
\newcommand{\pols}{\mathcal{P}}
\newcommand{\cants}{\Delta\!\!\!\!\Delta}
\newcommand{\eps}{\varepsilon}
\newcommand{\st}{\backepsilon}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\dom}[1]{\mathrm{dom}\left(#1\right)}
\newcommand{\for}{\text{ for }}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\spn}{\mathrm{sp}}
\newcommand{\nul}{\mathcal{N}}
\newcommand{\col}{\mathrm{col}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\norm}[1]{\lVert #1 \rVert}
\newcommand{\inner}[1]{\left\langle #1 \right\rangle}
\newcommand{\pmat}[1]{\begin{pmatrix} #1 \end{pmatrix}}
\renewcommand{\and}{\text{ and }}

\newsavebox{\qed}
\newenvironment{proof}[2][$\square$]
    {\setlength{\parskip}{0pt}\par\textit{Proof:} #2\setlength{\parskip}{0.25cm}
        \savebox{\qed}{#1}
        \begin{adjustwidth}{\widthof{Proof:}}{}
    }
    {
        \hfill\usebox{\qed}\end{adjustwidth}
    }

\pagestyle{fancy}
\fancyhead{}
\lhead{Caleb Jacobs}
\chead{APPM 5610: Numerical Analysis II}
\rhead{Homework \#4}
\cfoot{}
\setlength{\headheight}{35pt}
\setlength{\parskip}{0.25cm}
\setlength{\parindent}{0pt}

\begin{document}
\begin{enumerate}[label = (\arabic*)]
	\item Show that Jacobi's method for finding eigenvalues of a real symmetric matrix is ultimately quadratically convergent. Assume that all off-diagonal elements of the matrix $ A_k $ are $ O(\eps) $,where $ k $ enumerates Jacobi sweeps. Show that the all rotations of the next Jacobi sweep are of the form
	\[
		\pmat{
			1 & 0 & \cdots & \cdots & \cdots & \cdots & 0 \\
			0 & 1 & \cdots & \cdots & \cdots & \cdots & 0 \\
			\cdots & \cdots & \cdots & \cdots & \cdots & \cdots & \cdots \\
			0 & \cdots & 1- O(\eps^2) & \cdots & O(\eps) & \cdots & 0 \\
			\cdots & \cdots & \cdots & \cdots & \cdots & \cdots & \cdots \\
			0 & \cdots & O(\eps) & \cdots & 1 - O(\eps^2) & \cdots & 0 \\
			0 & \cdots & \cdots & \cdots & \cdots & 1 & 0 \\
			0 & \cdots & \cdots & \cdots & \cdots & 0 & 1
		}
	\]
	Then demonstrate that this implies that, after the sweep, all off-diagonal elements of $ A_{k+1} $ are $ O(\eps^2) $. Assume that all eigenvalues are non-zero and distinct.
	
	\newpage
	\item Show that $ A $ is diagonalizable iff there is a positive definite self-adjoint matrix $ H $ such that $ H^{-1} A H $ is normal.
	
	\begin{proof}{}
		($ \implies $)Suppose a matrix $ A $ is diagonalizable. Then we can write
		\[
			A = P D P^{-1} \implies D = P^{-1} A P
		\]
		where $ D $ is a diagonal and $ P $ is an invertible. Next, let's take the polar decomposition of $ P $ as $ H = HU $ where $ H $ is positive definite self-adjoint and $ U $ is unitary. Note, $ H^{-1} $ is also positive definite-self-adjoint and $ H = PU^* $. Then
		\[
			H^{-1} A H = U P^{-1} A P U^* = U D U^*.
		\]
		Then
		\begin{align*}
			(H^* A H)^*(H^* A H) &= (U D U^*)^*(U D U^*) \\
			&= (U D^* U^*)(U D U^*) \\
			&= U D^* D U^* \\
			&= U D D^* U^* \\
			&= (U D U^*) (U D^* U^*) \\
			&= (U D U^*) (U D U^*)^* \\
			&= (H^{-1} A H) (H^{-1} A H)^*
		\end{align*}
		Showing that there exists a positive definite self-adjoint matrix $ H  $ such that $ H^{-1}AH $ is normal.
		
		($ \impliedby $) Now, suppose there exists a positive definite self-adjoint matrix $ H $ such that $ H^{-1} A H $ is normal. Then, because $ H^{-1} A H $ is normal, it is diagonalizable by a unitary matrix:
		\[
			H^{-1} A H = U D U^*
		\]
		where $ D $ is diagonal and $ U $ is unitary. Then, we have
		\[
			D = U^* H^{-1} A H U = P^{-1} A P 
		\]
		where $ P = HU $. So, $ A $ is similar to a diagonal matrix and is thus diagonalizable.
	\end{proof}
\end{enumerate}
\end{document}